---
title: "How Political Media Exposure Influences Voter Turnout Across Age Groups in US Elections"
author: Anna Ellison and William Lewis
format: revealjs
editor: visual
editor_options: 
  chunk_output_type: inline
---
## Introduction

Topic: How Political Media Exposure Influences Voter Turnout Across Age Groups in US Elections

- Our project aims to investigate how the volume and type of political media, especially social media posts and advertisements, correlate with voter turnout across different age demographics
- We specifically plan to examine the rise in political advertisements and social media posts surrounding election dates, and how this can ultimately impact voter turnout and appeal to different generations of media users

```{r}
#| include: FALSE
library(tidyverse)
library(janitor)
library(data.table)
library(dplyr)
library(ggplot2)
library(usmap)
library(scales)
library(tinytable)
library(maps)
library(gt)
library(RColorBrewer)
```

```{r}
tweetvolume <- read.csv("tweet_volume_data_refined.csv")
tweetvolume$date <- as.Date(tweetvolume$date)
```

## Data from X
```{r}
ggplot(tweetvolume, aes(x = date, y = tweet_count)) +
  geom_area(fill = "#B3CDE0", alpha = 0.4) +
  geom_line(color = "005B96", linewidth = 1.5) +
  geom_point(color = "#03396C", size = 2) +
  scale_x_date(date_labels = waiver(), date_breaks = waiver()) +
  scale_y_continuous(labels = comma) +
  labs(
      title = "Social Media Surge: Political Tweets Leading to the 2024 Election",
      subtitle = "Poltical Tweet volume increases as election day approaches",
      x = "Date",
      y = "Number of Tweets",
      caption = "Source: X"
  ) +
theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13, color = "gray40"),
    axis.title = element_text(face = "bold"),
    axix.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )
```

```{r}
voterturnout2016 <- read.csv("voterturnout2016.csv") |>
  group_by(state) |>
  rename('Voter Turnout Rate 2016 (18-24)' = Share.of.Pop.Ages.18...24) |>
  rename('Voter Turnout Rate 2016 (45-64)' = Share.of.Pop.Ages.45...64)

youngvoters2016 <- voterturnout2016 |>
  select('Voter Turnout Rate 2016 (18-24)') 

youngvoters2016$`Voter Turnout Rate 2016 (18-24)` <- as.numeric(youngvoters2016$`Voter Turnout Rate 2016 (18-24)`)

youngvoters2016$`Voter Turnout Rate 2016 (18-24)` <- youngvoters2016$`Voter Turnout Rate 2016 (18-24)` * 100

oldervoters2016 <- voterturnout2016 |>
  select('Voter Turnout Rate 2016 (45-64)')

oldervoters2016$'Voter Turnout Rate 2016 (45-64)' <- as.numeric(oldervoters2016$'Voter Turnout Rate 2016 (45-64)')

oldervoters2016$'Voter Turnout Rate 2016 (45-64)' <- oldervoters2016$'Voter Turnout Rate 2016 (45-64)' * 100
```

## Voter Turnout 2016 Heat Map
```{r}
plot_usmap(data = youngvoters2016, values = "Voter Turnout Rate 2016 (18-24)", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90", limits = c(20,100)) +
  labs(title = "Voter Turnout Rate (Ages 18-24) by State in 2016") +
  theme(plot.title = element_text(size = 18, face = "bold"),
    legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```


```{r}
voterturnout2020 <- read.csv("voter_turnout.csv") |>
    group_by(State) |>
  rename(state = State) |>
  rename("Voter Turnout Rate 18-24" = Share.of.Pop..18...24) |>
  rename("Voter Turnout Rate 45-64" = Share.of.Pop..45...64)

youngvoters2020 <- voterturnout2020 |>
  select(`Voter Turnout Rate 18-24`)

youngvoters2020$`Voter Turnout Rate 18-24` <- as.numeric(youngvoters2020$`Voter Turnout Rate 18-24`)

youngvoters2020$`Voter Turnout Rate 18-24` <- youngvoters2020$`Voter Turnout Rate 18-24` * 100

oldervoters2020 <- voterturnout2020 |>
  select('Voter Turnout Rate 45-64')

oldervoters2020$'Voter Turnout Rate 45-64' <- as.numeric(oldervoters2020$'Voter Turnout Rate 45-64')

oldervoters2020$'Voter Turnout Rate 45-64' <- oldervoters2020$'Voter Turnout Rate 45-64' * 100

```

## Voter Turnout 2020 Heat Map
```{r echo = FALSE}
plot_usmap(data = youngvoters2020, values = "Voter Turnout Rate 18-24", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90", limits = c(20,100)) +
  labs(title = "Voter Turnout Rate (Ages 18-24) by State in 2020") +
  theme(plot.title = element_text(size = 18, face = "bold"), legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```

## Older Voter Turnout 2016 Heat Map
```{r}
plot_usmap(data = oldervoters2016, values = 'Voter Turnout Rate 2016 (45-64)', regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90", limits = c(52,100)) +
  labs(title = "Voter Turnout Rate (Ages 45-64) by State in 2016") +
  theme(plot.title = element_text(size = 18, face = "bold"),
    legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```

## Older Voter Turnout 2020 Heat Map
```{r echo = FALSE}
plot_usmap(data = oldervoters2020, values = "Voter Turnout Rate 45-64", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90", limits = c(52,100)) +
  labs(title = "Voter Turnout Rate (Ages 45-64) by State in 2020") +
  theme(plot.title = element_text(size = 18, face = "bold"), legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```


```{r}
geo_spend <- read.csv("google-political-ads-advertiser-geo-spend.csv") 

statespending <- geo_spend |>
  filter(Country == "US") |>
  rename(state = Country_Subdivision_Primary) |>
  group_by(state) |>
  filter(state != "AS", state != "GU", state != "MP", state != "DC", state != "PR", state != "VI" ,state != "") |>
  summarize(Total = sum(Spend_USD)/1000000)
```

```{r}
statespendingpercapita <- geo_spend |>
  filter(Country == "US") |>
  rename(state = Country_Subdivision_Primary) |>
  group_by(state) |>
  filter(state != "AS", state != "GU", state != "MP", state != "DC", state != "PR", state != "VI" ,state != "") |>
  select("state", "Spend_USD") |>
  summarize(Total = sum(Spend_USD))
```

```{r}
statepopulations <- read.csv("CENSUSpopulationdata.csv") |>
  select("NAME", "POPESTIMATE2024") |>
  rename(state = NAME) |>
  filter(state != "United States", state != "Northeast Region", state != "New England", state != "Middle Atlantic", state != "Midwest Region", state != "East North Central", state != "West North Central", state != "South Region", state != "South Atlantic", state != "East South Central", state != "West South Central", state != "West Region", state != "Mountain", state != "Pacific", state != "Puerto Rico")

statepopulations <- statepopulations |>
  mutate(state = state.abb[match(state, state.name)])

mergedspendingpercapita <- merge(statespendingpercapita, statepopulations, by = "state")

mergedspendingpercapita <- mergedspendingpercapita |>
  mutate('Spending Per Capita' = Total/POPESTIMATE2024)
```

```{r include = FALSE}
plot_usmap(data = statespending, values = "Total", regions = "states") +
  scale_fill_continuous( low = "light yellow", high = "dark red", name = "Spending (in millions)", limits = c(2,225)) +
  theme(legend.position = "right", plot.title = element_text(size = 20, face = "bold")) +
  labs(title = "Heat Map of Political Ad Spending (Google) per State")
```

## Spending Per Capita Heat Map
```{r}
plot_usmap(data = mergedspendingpercapita, values = "Spending Per Capita", regions = "states") +
  scale_fill_continuous( low = "lightblue", high = "darkblue", name = "Spending Per Capita", limits = c(1,25)) +
  theme(legend.position = "right", plot.title = element_text(size = 20, face = "bold")) +
  labs(title = "Per Capita Political Ad Spending (Google) per State")
```

## Top 10 States With Highest Political Ad Spending (Per Capita) on Google
```{r echo = FALSE}
top10spendingstates <- mergedspendingpercapita |>
  select(state, `Spending Per Capita`)|>
  arrange(desc(`Spending Per Capita`)) |>
  slice_head(n = 10) |>
  rename(State = state)
  
tt(top10spendingstates) |>
  style_tt(fontsize = 0.72)
```

```{r}
mergedoldervoters <- merge(oldervoters2016, oldervoters2020, by = "state") |>
  mutate("Change in Voter Turnout Rate (45-64)" = `Voter Turnout Rate 45-64` - `Voter Turnout Rate 2016 (45-64)`)

top10changeinoldervoter <- mergedoldervoters |>
  select(state,"Change in Voter Turnout Rate (45-64)") |>
  arrange(desc(`Change in Voter Turnout Rate (45-64)`)) |>
  head(10) |>
  rename("State" = state)
```

## Top 10 States With Highest Increase in Voter Turnout Rates 2016-2020
```{r}
tt(top10changeinoldervoter) |>
  style_tt(fontsize = 0.72)
```

```{r}
facebookadlocations <- read.csv("facebookadlocations.csv") |>
  rename(state = "Location.name") |>
  rename(Spending = "Amount.spent..USD.") |>
  filter(state != "American Samoa", state != "Guam", state != "Northern Mariana Islands", state != "Puerto Rico", state != "Unknown", state != "Virgin Islands (U.S.)") |>
  mutate(Spending = Spending/1000000 )
```

```{r}
mergedyoungervoters <- merge(youngvoters2016, youngvoters2020, by = "state") |>
  mutate("Change in Voter Turnout Rate (18-24)" = `Voter Turnout Rate 18-24` - `Voter Turnout Rate 2016 (18-24)`)

top10changeinyoungervoter <- mergedyoungervoters |>
  select(state,"Change in Voter Turnout Rate (18-24)") |>
  arrange(desc(`Change in Voter Turnout Rate (18-24)`)) |>
  head(10) |>
  rename("State" = state)
```

## Top 10 States With Highest Increase in Voter Turnout Rates 2016-2020
```{r}
tt(top10changeinyoungervoter) |>
  style_tt(fontsize = 0.72)
```

```{r include = FALSE}
plot_usmap(data = facebookadlocations, values = "Spending", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Spending (in millions)") +
  theme(
    legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```

