---
title: "How Political Media Exposure Influences Voter Turnout Across Age Groups in US Elections"
author: Anna Ellison and William Lewis
format: revealjs
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Introduction

Topic: How Political Media Exposure Influences Voter Turnout Across Age Groups in US Elections

- Our project aims to investigate how the volume and type of political media, especially social media posts and advertisements, correlate with voter turnout across different age demographics
- We specifically plan to examine the rise in political advertisements and social media posts surrounding election dates, and how this can ultimately impact voter turnout and appeal to different generations of media users

```{r}
#| include: FALSE
library(tidyverse)
library(janitor)
library(data.table)
library(dplyr)
library(ggplot2)
library(usmap)
library(scales)
library(tinytable)
library(maps)
library(gt)
library(RColorBrewer)
```

```{r}
tweetvolume <- read.csv("tweet_volume_data_refined.csv")
tweetvolume$date <- as.Date(tweetvolume$date)
```

## Data from X
```{r}
ggplot(tweetvolume, aes(x = date, y = tweet_count)) +
  geom_area(fill = "#B3CDE0", alpha = 0.4) +
  geom_line(color = "005B96", linewidth = 1.5) +
  geom_point(color = "#03396C", size = 2) +
  scale_x_date(date_labels = waiver(), date_breaks = waiver()) +
  scale_y_continuous(labels = comma) +
  labs(
      title = "Social Media Surge: Tweets Leading to the 2024 Election",
      subtitle = "Tweet volume increases as election day approaches",
      x = "Date",
      y = "Number of Tweets",
      caption = "Source: X"
  ) +
theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13, color = "gray40"),
    axis.title = element_text(face = "bold"),
    axix.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )
```

```{r}
geo_spend <- read.csv("google-political-ads-advertiser-geo-spend.csv") 

statespending <- geo_spend |>
  filter(Country == "US") |>
  rename(state = Country_Subdivision_Primary) |>
  group_by(state) |>
  filter(state != "AS", state != "GU", state != "MP", state != "DC", state != "PR", state != "VI" ,state != "") |>
  summarize(Total = sum(Spend_USD)/1000000)
```

## Spending Heat Map
```{r echo = FALSE}
plot_usmap(data = statespending, values = "Total", regions = "states") +
  scale_fill_continuous( low = "light yellow", high = "dark red", name = "Spending (in millions)") +
  theme(legend.position = "right", plot.title = element_text(size = 20, face = "bold")) +
  labs(title = "Heat Map of Political Ad Spending per State")
```


```{r}
voterturnout2020 <- read.csv("voter_turnout.csv") |>
    group_by(State) |>
  rename(state = State) |>
  rename("Voter Turnout Rate" = Share.of.Pop..18...24)

youngvoters2020 <- voterturnout2020 |>
  select(`Voter Turnout Rate`) |>
  filter(`Voter Turnout Rate` != "NSD", `Voter Turnout Rate` != 0)

youngvoters2020$`Voter Turnout Rate` <- as.numeric(youngvoters2020$`Voter Turnout Rate`)

youngvoters2020$`Voter Turnout Rate` <- youngvoters2020$`Voter Turnout Rate` * 100

```

## Voter Turnout 2020 Heat Map
```{r echo = FALSE}
plot_usmap(data = youngvoters2020, values = "Voter Turnout Rate", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90") +
  labs(title = "Voter Turnout Rate (Ages 18-24) by State in 2020") +
  theme(plot.title = element_text(size = 18, face = "bold"), legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```

```{r}
voterturnout2016 <- read.csv("voterturnout2016.csv") |>
  group_by(state) |>
  rename("Voter Turnout Rate 2016" = Share.of.Pop.Ages.18...24)

youngvoters2016 <- voterturnout2016 |>
  select('Voter Turnout Rate 2016') |>
  filter(`Voter Turnout Rate 2016` != "NSD", `Voter Turnout Rate 2016` != 0)

youngvoters2016$`Voter Turnout Rate 2016` <- as.numeric(youngvoters2016$`Voter Turnout Rate 2016`)

youngvoters2016$`Voter Turnout Rate 2016` <- youngvoters2016$`Voter Turnout Rate 2016` * 100
```

## Voter Turnout 2016 Heat Map
```{r}
plot_usmap(data = youngvoters2016, values = "Voter Turnout Rate 2016", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Turnout Rate", na.value = "gray90") +
  labs(title = "Voter Turnout Rate (Ages 18-24) by State in 2016") +
  theme(plot.title = element_text(size = 18, face = "bold"),
    legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```


## Top 10 States With Highest Voter Turnout Rate Ages 18 - 24
```{r echo = FALSE}
top10youngstates2020 <- youngvoters2020 |>
  arrange(desc(`Voter Turnout Rate`)) |>
  head(10) |>
  rename(State = "state")
```

```{r}
tt(top10youngstates2020) |>
  style_tt(fontsize = 0.72)
```
## Top 10 States With Spending On Political Advertisements On Google
```{r echo = FALSE}
top10spendingstates <- statespending |>
  arrange(desc(Total)) |>
  head(10) |>
  rename(State = "state")
  
tt(top10spendingstates) |>
  style_tt(fontsize = 0.72)
```

```{r}
facebookadlocations <- read.csv("facebookadlocations.csv") |>
  rename(state = "Location.name") |>
  rename(Spending = "Amount.spent..USD.") |>
  filter(state != "American Samoa", state != "Guam", state != "Northern Mariana Islands", state != "Puerto Rico", state != "Unknown", state != "Virgin Islands (U.S.)") |>
  mutate(Spending = Spending/1000000 )
```

## Facebook Political Ad Spending
```{r echo = FALSE}
plot_usmap(data = facebookadlocations, values = "Spending", regions = "states") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", name = "Spending (in millions)") +
  theme(
    legend.background = element_rect(fill = "transparent", color = NA),
      legend.position = c(0,0.05))
```


