---
title: "Impact of Media Forms on Voting Preference and Behavior"
author: "Anna Ellison and William Lewis"
format: revealjs
---

## Introduction

- Our project aims to investigate the impact of various media forms on voting turnout and behaviors.
- We plan to use publicly available datasets to examine correlations between different sources of media consumption and political preferences and behavior. 
- The primary goal of our project is to find statiscally significant data to help provide an understanding of media influence and democratic participation. 

```{r}
#| include: FALSE
library(tidyverse)
library(janitor)
library(data.table)
library(dplyr)
library(ggplot2)
```

```{r}
#| include: FALSE
anesspecialstudy <- read.csv("anes_specialstudy_2020-2022_socialmedia_csv_20230705.csv") 
```

```{r}
#| include: FALSE
newsonline <- anesspecialstudy |>
  select(profile_newsonline, profile_newsmobile, trustfox, trustfb, trusttw, trustmsnbc)

newsonline <- newsonline %>%
  mutate(newsyesterday = case_when(profile_newsonline == 1 ~ "Yes", profile_newsonline == 2 ~ "No")) |>
  mutate(foxtrust = case_when(trustfox == 1 ~ "Not at all", trustfox == 2 ~ "A little", trustfox == 3 ~ "A moderate amount", trustfox == 4 ~ "A lot", trustfox == 5 ~ "A great deal")) |>
  mutate(msnbctrust = case_when(trustmsnbc == 1 ~ "Not at all", trustmsnbc == 2 ~ "A little", trustmsnbc == 3 ~ "A moderate amount", trustmsnbc == 4 ~ "A lot", trustmsnbc == 5 ~ "A great deal")) |>
  mutate(fbtrust = case_when(trustfb == 1 ~ "Not at all", trustfb == 2 ~ "A little", trustfb == 3 ~ "A moderate amount", trustfb == 4 ~ "A lot", trustfb == 5 ~ "A great deal"))
```

```{r}
#| include: FALSE
# The ANES (American National Elections Study) 2020-2022 Social Media Study is a three-wave panel survey conducted on the Internet to provide data about voting and public opinion in the 2020 and 2022 elections. Over 5500 people were interviewed for this survey. 
```

## Use of Online News
```{r}
frequency_table <- table(newsonline$newsyesterday)

barplot(frequency_table, main = "Did you get any news online yesterday?", xlab = "", ylab = "Frequency")
```

## Trust in Media Forms
```{r}
fbtrustfrequency <- table(newsonline$fbtrust)

barplot(fbtrustfrequency, main = "How trustworthy do you believe Facebook to be?", xlab = "", ylab = "Frequency")
```

